---
title: "Scenario IV on real data: per-drop and joint analyses"
author: ""
date: ""
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
    number_sections: true
    df_print: paged
    fig_width: 13
    fig_height: 8
    code_folding: hide
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  echo       = TRUE,
  warning    = FALSE,
  message    = FALSE,
  fig.align  = "center",
  cache      = FALSE,
  results    = "asis"
)
```


```{r project-setup}
# Common setup for the whole report
project_root <- "/Users/negar/Documents/phd/estimate_mixture_models-main"
knitr::opts_knit$set(root.dir = project_root)
```

```{r libraries}
library(truncnorm)
library(readxl)
library(ggplot2)
library(gridExtra)
library(patchwork)
library(tidyr)
library(MASS)
library(mvtnorm)
library(invgamma)
```

```{r sources}
source(file.path(project_root, "scripts/physics_model.R"))
source(file.path(project_root, "scripts/helper_function_CGP.R"))
source(file.path(project_root, "scripts/main_function_seuil_CGP.R"))
```

# Part 1 — Scenario IV per drop, all sheets


```{r per-drop-config}
# ---------------------------------------------------------------------
# Output directory for the per-(sheet, drop) result files
# ---------------------------------------------------------------------
out_dir <- file.path(project_root, "results/real_data_seuil_gfix")
if (!dir.exists(out_dir)) dir.create(out_dir, recursive = TRUE)

# ---------------------------------------------------------------------
# List of sheets in the Excel file (sheet index -> ball name)
# ---------------------------------------------------------------------
ball_sheets <- list(
  list(idx = 1,  name = "Baseball"),
  list(idx = 2,  name = "Blue_Basketball"),
  list(idx = 3,  name = "Green_Basketball"),
  list(idx = 4,  name = "Volleyball"),
  list(idx = 5,  name = "Bowling_Ball"),
  list(idx = 6,  name = "Golf_Ball"),
  list(idx = 7,  name = "Tennis_Ball"),
  list(idx = 8,  name = "Whiffle_Ball_1"),
  list(idx = 9,  name = "Whiffle_Ball_2"),
  list(idx = 10, name = "Yellow_Whiffle_Ball"),
  list(idx = 11, name = "Orange_Whiffle_Ball")
)

# ---------------------------------------------------------------------
# MCMC settings (identical to examples/real_data.R, do not change)
# ---------------------------------------------------------------------
n_iter      <- 10000
burn_in     <- 2000
sigma_props <- c(NA, NA, NA, NA, 0.5, NA)
# init = c(g, h0, sig2err, alpha, psidelta, k)
init_template <- c(9.8, NA, 0.01, 0.5, 0.5, 0.1)

xlsx_path <- file.path(project_root, "data", "Ball_drops_data.xlsx")
stopifnot(file.exists(xlsx_path))
```

```{r per-drop-helpers}
# Composite figure for one MCMC result (same layout as in
# plot_real_data_one_result.R)
plot_composite_one <- function(result_one) {

  ball <- result_one$ball
  dnum <- result_one$drop
  n    <- result_one$n
  res  <- result_one$res

  theta_mat <- res[["theta"]]
  delta_mat <- res[["delta"]]
  zeta_mat  <- res[["zeta"]]

  g     <- theta_mat[, 1]
  h0    <- theta_mat[, 2]
  sigma <- theta_mat[, 3]
  alpha <- theta_mat[, 4]
  psi   <- theta_mat[, 5]
  k     <- theta_mat[, 6]

  # alpha density
  df_alpha <- data.frame(alpha = alpha)
  p_alpha <- ggplot(df_alpha, aes(x = alpha)) +
    geom_density(fill = "#4CCDC9", color = "lightseagreen",
                 alpha = 0.6, linewidth = 1) +
    labs(x = expression(alpha), y = "Density",
         title = paste0("Posterior density of alpha (real data, ",
                        ball, ", drop ", dnum, ")")) +
    theme_minimal(base_size = 13) +
    theme(plot.title = element_text(hjust = 0.5))

  # delta(x_i) summary
  delta_mean <- colMeans(delta_mat)
  delta_ci   <- t(apply(delta_mat, 2, quantile, probs = c(0.025, 0.975)))
  df_delta <- data.frame(
    i = 1:n,
    d_mean = delta_mean,
    d_low  = delta_ci[, 1],
    d_high = delta_ci[, 2]
  )
  p_delta <- ggplot(df_delta, aes(x = i, y = d_mean)) +
    geom_ribbon(aes(ymin = d_low, ymax = d_high),
                fill = "lightseagreen", alpha = 0.25) +
    geom_line(color = "lightseagreen", linewidth = 0.9) +
    geom_hline(yintercept = 0, linetype = "dashed") +
    labs(x = "Index i", y = expression(delta(x[i])),
         title = paste0("Posterior summary of delta(x_i) (",
                        ball, ", drop ", dnum, ")")) +
    theme_minimal(base_size = 12) +
    theme(plot.title = element_text(hjust = 0.5))

  # pointwise inclusion probabilities
  p_hat <- colMeans(zeta_mat == 2)
  df_p  <- data.frame(i = 1:n, p_hat = p_hat)
  p_zeta <- ggplot(df_p, aes(x = i, y = p_hat)) +
    geom_col(fill = "lightseagreen") +
    ylim(0, 1) +
    labs(x = "Index i",
         y = expression(hat(p)[i] == P(zeta[i] == 1 ~ "|" ~ Y)),
         title = paste0("Pointwise inclusion probabilities (",
                        ball, ", drop ", dnum, ")")) +
    theme_minimal(base_size = 12) +
    theme(plot.title = element_text(hjust = 0.5))

  # boxplot grid
  make_box <- function(x, ylab) {
    df <- data.frame(val = x)
    ggplot(df, aes(x = 1, y = val)) +
      geom_boxplot(fill = "lightseagreen", width = 0.25, outlier.size = 1) +
      labs(x = NULL, y = ylab) +
      theme_minimal(base_size = 11) +
      theme(axis.text.x  = element_blank(),
            axis.ticks.x = element_blank(),
            plot.margin  = margin(3, 6, 3, 6))
  }
  p_g     <- make_box(g,     ylab = "g")
  p_h0    <- make_box(h0,    ylab = "h0")
  p_sig   <- make_box(sigma, ylab = expression(lambda^2))
  p_a_box <- make_box(alpha, ylab = expression(alpha))
  p_psi   <- make_box(psi,   ylab = expression(gamma[delta]))
  p_k     <- make_box(k,     ylab = "k")

  left_grid <- (p_g | p_h0) /
               (p_sig | p_a_box) /
               (p_psi | p_k)

  right_col <- p_alpha / p_delta / p_zeta +
               patchwork::plot_layout(heights = c(1, 1.25, 1))

  fig <- left_grid | right_col
  fig <- fig + patchwork::plot_layout(widths = c(1.05, 1.45))
  fig
}

# Numerical posterior summary table
posterior_summary_table <- function(theta_mat) {
  summ <- function(x) c(mean = mean(x), median = median(x), sd = sd(x),
                        q025 = unname(quantile(x, 0.025)),
                        q975 = unname(quantile(x, 0.975)))
  tab <- rbind(
    g     = summ(theta_mat[, 1]),
    h0    = summ(theta_mat[, 2]),
    sigma = summ(theta_mat[, 3]),
    alpha = summ(theta_mat[, 4]),
    psi   = summ(theta_mat[, 5]),
    k     = summ(theta_mat[, 6])
  )
  round(tab, 4)
}
```

```{r per-drop-runner, results = "asis"}

for (sh in ball_sheets) {

  don <- read_xlsx(xlsx_path, sheet = sh$idx)
  names(don) <- c("drop", "time", "Height", "Velocity")
  don$drop <- as.factor(don$drop)
  drops_in_sheet <- levels(don$drop)

  for (dnum in drops_in_sheet) {

    sub <- don[don$drop == dnum, ]
    if (nrow(sub) < 3) next

    # ----- Per-drop preprocessing (mirrors data/prepare_data.R) -----
    t       <- sub$time
    y       <- sub$Height
    t_min   <- min(t)
    t_range <- max(t) - min(t)
    t       <- (t - t_min) / t_range
    n       <- length(y)
    a       <- t_range

    # ----- Per-drop initial height ---------------------------------
    init    <- init_template
    init[2] <- max(y)

    cat("\n\n## ", sh$name, " — drop ", dnum, "\n\n", sep = "")
    cat("- `n        = `", n, "\n")
    cat("- `t_range  = `", round(t_range, 4), " s\n")
    cat("- `init[2]  = h0 = `", round(init[2], 4), "\n\n")

    # ----- Reproducible chain ---------------------------------------
    set.seed(12345)

    res <- mcmc_step6(
      y               = y,
      t               = t,
      n_iter          = n_iter,
      init            = init,
      sigma_proposals = sigma_props,
      g_init          = TRUE,
      h0_init         = FALSE,
      sig2er_init     = FALSE,
      alpha_init      = FALSE,
      psi_init        = FALSE,
      k_init          = FALSE,
      Sigma_theta     = matrix(c(0.5, 0, 0, 0.5), 2),
      n_burnin        = burn_in,
      seuil           = TRUE,
      s               = 0.3
    )

    result_one <- list(
      sheet_idx = sh$idx,
      ball      = sh$name,
      drop      = dnum,
      n         = n,
      t         = t,
      y         = y,
      t_min     = t_min,
      t_range   = t_range,
      init      = init,
      res       = res
    )

    # ----- Save -----------------------------------------------------
    out_file <- file.path(
      out_dir,
      sprintf("real_seuil_gfix_%s_drop%s.RData", sh$name, dnum)
    )
    save(result_one, file = out_file)

    # ----- Posterior summary as an HTML table -----------------------
    cat("\n\n**Posterior summary**\n\n")
    print(knitr::kable(posterior_summary_table(res$theta),
                       format  = "html",
                       caption = paste0("Posterior summary — ", sh$name,
                                        ", drop ", dnum)))

    # ----- Composite figure -----------------------------------------
    print(plot_composite_one(result_one))

    cat("\n\n")
  }
}
```

# Part 2 — Joint analysis on Blue Basketball

In the **joint analysis**, we build **one** dataset by combining the two
drops of Blue Basketball, and we run **one** MCMC on it. The model is
slightly extended so that:

- there are **two separate** intercepts $h_0^{(1)}$ and $h_0^{(2)}$
  (each drop is released from a slightly different height),

The two drops are the same physical experiment repeated
twice, so they should share the same drag function. Pooling the two
datasets gives more observations to estimate $\delta(\cdot)$.

1. `init` has length 7 instead of 6, with two intercepts `h0_1` and
   `h0_2` instead of one `h0`.
2. The Gaussian-process covariance matrix $\Sigma_\delta$ is built on
   the joint normalized time vector (the union of both drops' times),
   so it is $N \times N$ with $N = n_1 + n_2$ instead of $n \times n$.
3. The Gibbs step that samples the calibration parameters now updates
   $(h_0^{(1)}, h_0^{(2)}, g)$ jointly (a 3-D Gaussian) instead of
   $(h_0, g)$ (a 2-D Gaussian).

## MCMC


```{r joint-mcmc-function}
mcmc_step6_joint <- function(y, t, d_idx, n_iter, init, sigma_proposals,
                             g_init   = TRUE,
                             h0_init  = TRUE,
                             sig2er_init = TRUE,
                             alpha_init  = TRUE,
                             psi_init    = TRUE,
                             k_init      = TRUE,
                             Sigma_theta,
                             n_burnin = 1000,
                             seuil = FALSE, s = 0.3) {

  total_iter <- n_burnin + n_iter

  N <- length(y)
  stopifnot(length(t) == N, length(d_idx) == N, all(d_idx %in% c(1L, 2L)))

  # init = c(g, h0_1, h0_2, sigma_sq_err, alpha, psi_delta, k)
  theta <- init
  delta <- rep(0, N)

  chain_theta <- matrix(NA, nrow = total_iter, ncol = 7)
  colnames(chain_theta) <- c("g", "h0_1", "h0_2", "sigma_sq_err",
                             "alpha", "psi_delta", "k")
  chain_delta <- matrix(NA, nrow = total_iter, ncol = N)
  chain_zeta  <- matrix(NA, nrow = total_iter, ncol = N)

  loglik_chain <- numeric(total_iter)
  accept_psi   <- 0

  # Drop indicators as a 0/1 design matrix (N x 2)
  D <- cbind(as.numeric(d_idx == 1), as.numeric(d_idx == 2))

  for (iter in 1:total_iter) {
    g            <- theta[1]
    h0_1         <- theta[2]
    h0_2         <- theta[3]
    sigma_sq_err <- theta[4]
    alpha_param  <- theta[5]
    psi_delta    <- theta[6]
    k            <- theta[7]

    sigma_sq_delta <- sigma_sq_err / k

    Sigma_delta <- GP_covariance(t, sigma_sq_delta, psi_delta)

    # Per-observation mean: h0_d - 0.5*g*t^2  (vector of length N)
    h0_vec  <- ifelse(d_idx == 1, h0_1, h0_2)
    f_theta <- h0_vec - 0.5 * g * (t * t_range + t_min)^2
    f_theta[f_theta < 0] <- 0   

    mean1 <- f_theta
    mean2 <- f_theta + delta

    # ---------------- zeta update (log-sum-exp, same as mcmc_step6) -
    log_w1 <- log(alpha_param) + dnorm(y, mean1, sqrt(sigma_sq_err), log = TRUE)
    log_w2 <- log(1 - alpha_param) + dnorm(y, mean2, sqrt(sigma_sq_err), log = TRUE)
    log_max <- pmax(log_w1, log_w2)
    log_den <- log_max + log(exp(log_w1 - log_max) + exp(log_w2 - log_max))

    if (seuil) {
      prob_zeta_base <- exp(log_w2 - log_den)
      prob_zeta      <- prob_zeta_base * (abs(delta) > s)
    } else {
      prob_zeta <- exp(log_w2 - log_den)
    }
    zeta <- ifelse(runif(N) < prob_zeta, 2, 1)
    chain_zeta[iter, ] <- zeta

    # ---------------- log-likelihood ------------------------------
    log_likelihood <- sum(log(ifelse(zeta == 1, alpha_param * dnorm(y, mean1, sqrt(sigma_sq_err)), (1 - alpha_param) * dnorm(y, mean2, sqrt(sigma_sq_err)))))
    loglik_chain[iter] <- log_likelihood

    # ---------------- delta --------------------------------
    zeta_2_indices <- which(zeta == 2)
    if (length(zeta_2_indices) > 0) {
      y_m              <- y[zeta_2_indices]
      Sigma_delta_ymym <- sigma_sq_err * diag(length(zeta_2_indices)) +
        Sigma_delta[zeta_2_indices, zeta_2_indices, drop = FALSE]
      Sigma_delta_ym   <- Sigma_delta[, zeta_2_indices, drop = FALSE]
      Sigma_inv        <- tryCatch(solve(Sigma_delta_ymym),
                                   error = function(e) diag(1, nrow(Sigma_delta_ymym)))
      mu_delta_hat     <- rep(0, N) +
        Sigma_delta_ym %*% Sigma_inv %*%
        (y_m - f_theta[zeta_2_indices])
      Sigma_delta_hat  <- Sigma_delta -
        Sigma_delta_ym %*% Sigma_inv %*% t(Sigma_delta_ym)
      Sigma_delta_hat  <- 0.5 * (Sigma_delta_hat + t(Sigma_delta_hat))
      delta <- as.vector(rmvnorm(1, mean = mu_delta_hat,
                                 sigma = Sigma_delta_hat))
    } else {
      delta <- as.vector(rmvnorm(1, rep(0, N), sigma = Sigma_delta))
    }

    # ---------------- Joint Gibbs for (h0_1, h0_2, g) -------
    zeta_1_indices <- which(zeta == 1)
    zeta_2_indices <- which(zeta == 2)

    Xfull <- cbind(D, -0.5 * (t * t_range)^2)         # N x 3
    x1 <- Xfull[zeta_1_indices, , drop = FALSE]
    x2 <- Xfull[zeta_2_indices, , drop = FALSE]

    y1 <- matrix(y[zeta_1_indices], ncol = 1)
    y2 <- matrix(y[zeta_2_indices], ncol = 1)
    d2 <- matrix(delta[zeta_2_indices], ncol = 1)

    A <- (t(x1) %*% x1 + t(x2) %*% x2) / sigma_sq_err
    B <- (t(x1) %*% y1 + t(x2) %*% y2 - t(x2) %*% d2) / sigma_sq_err

    Sigmapost_theta <- solve(A)
    Mupost_theta    <- Sigmapost_theta %*% B

    theta_sample <- rmvnorm(1, mean = Mupost_theta, sigma = Sigmapost_theta)
    h0_1_new <- theta_sample[1]
    h0_2_new <- theta_sample[2]
    g_new    <- theta_sample[3]

    if (g_init)  g_new  <- init[1]
    if (h0_init) {
      h0_1_new <- init[2]
      h0_2_new <- init[3]
    }

    g    <- g_new
    h0_1 <- h0_1_new
    h0_2 <- h0_2_new

    # ---------------- sigma_sq_err -------------------------
    if (sig2er_init) {
      sigma_sq_err <- init[4]
    } else {
      R <- outer(t, t, function(ti, tj) exp(-abs(ti - tj) / psi_delta))
      R_inv <- tryCatch(solve(R), error = function(e) diag(1, N))
      quad_form_delta <- as.numeric(t(delta) %*% R_inv %*% delta)

      h0_vec  <- ifelse(d_idx == 1, h0_1, h0_2)
      f_theta <- h0_vec - 0.5 * g * (t * t_range + t_min)^2
      f_theta[f_theta < 0] <- 0

      idx1 <- which(zeta == 1); idx2 <- which(zeta == 2)
      rss1 <- sum((y[idx1] - f_theta[idx1])^2)
      rss2 <- sum((y[idx2] - f_theta[idx2] - delta[idx2])^2)

      d <- 2
      rate_err  <- 0.5 * (rss1 + rss2 + k * quad_form_delta)
      shape_err <- N + d / 2
      sigma_sq_err <- rinvgamma(1, shape = shape_err, rate = rate_err)
    }

    # ---------------- psi_delta  ---
    if (psi_init) {
      psi_delta <- init[6]
    } else {
      psi_prop <- rtruncnorm(1, a = 0.1, b = 1,
                             mean = psi_delta, sd = sigma_proposals[6])
      sigma_sq_delta_prop <- sigma_sq_err / k
      Sigma_delta      <- GP_covariance(t, sigma_sq_delta_prop, psi_delta)
      Sigma_delta_prop <- GP_covariance(t, sigma_sq_delta_prop, psi_prop)

      log_prop_current  <- log(dtruncnorm(psi_delta, a = 0.1, b = 1,
                                          mean = psi_prop, sd = sigma_proposals[6]))
      log_prop_prop     <- log(dtruncnorm(psi_prop,  a = 0.1, b = 1,
                                          mean = psi_delta, sd = sigma_proposals[6]))
      log_prior_current <- dunif(psi_delta, 0.1, 1, log = TRUE)
      log_prior_prop    <- dunif(psi_prop,  0.1, 1, log = TRUE)

      log_like_current <- tryCatch(
        dmvnorm(delta, rep(0, N), Sigma_delta,      log = TRUE),
        error = function(e) -Inf)
      log_like_prop <- tryCatch(
        dmvnorm(delta, rep(0, N), Sigma_delta_prop, log = TRUE),
        error = function(e) -Inf)

      log_ratio <- (log_like_prop + log_prior_prop) -
        (log_like_current + log_prior_current) +
        (log_prop_current - log_prop_prop)

      if (!is.na(log_ratio) && log(runif(1)) < log_ratio) {
        psi_delta  <- psi_prop
        accept_psi <- accept_psi + 1
      }
    }

    # ---------------- k ----
    if (k_init) {
      k <- init[7]
    } else {
      alpha_k <- (N / 2) + 1
      beta_k  <- (1 / (2 * sigma_sq_err)) * quad_form_delta
      for (try_k in 1:100) {
        k_prop <- rgamma(1, shape = alpha_k, rate = beta_k)
        if (k_prop > 0 && k_prop < 1) { k <- k_prop; break }
      }
    }

    # ---------------- alpha --------------------------------
    if (alpha_init) {
      alpha_param <- init[5]
    } else {
      if (seuil) {
        zetabis <- zeta
        zetabis[abs(delta) < s] <- 0
        alpha_param <- rbeta(1, sum(zetabis == 1) + 0.5,
                             sum(zetabis == 2) + 0.5)
      } else {
        alpha_param <- rbeta(1, sum(zeta == 1) + 0.5,
                             sum(zeta == 2) + 0.5)
      }
    }

    theta <- c(g, h0_1, h0_2, sigma_sq_err, alpha_param, psi_delta, k)
    chain_theta[iter, ] <- theta
    chain_delta[iter, ] <- delta
  }

  if (n_burnin > 0) {
    keep <- (n_burnin + 1):total_iter
    chain_theta  <- chain_theta[keep, , drop = FALSE]
    chain_delta  <- chain_delta[keep, , drop = FALSE]
    chain_zeta   <- chain_zeta[keep, , drop = FALSE]
    loglik_chain <- loglik_chain[keep]
  }

  return(list(theta = chain_theta,
              delta = chain_delta,
              zeta  = chain_zeta,
              loglik = loglik_chain,
              accept_rate_psi = accept_psi / n_iter))
}
```

```{r joint-data}
out_dir_joint <- file.path(project_root, "results", "real_data_joint_seuil_gfix")
if (!dir.exists(out_dir_joint)) dir.create(out_dir_joint, recursive = TRUE)


don <- read_xlsx(xlsx_path, sheet = 2)
names(don) <- c("drop", "time", "Height", "Velocity")
don$drop <- as.factor(don$drop)

sub1 <- don[don$drop == 1, ]
sub2 <- don[don$drop == 2, ]

t_raw_1 <- sub1$time;  y_1 <- sub1$Height;  n_1 <- length(y_1)
t_raw_2 <- sub2$time;  y_2 <- sub2$Height;  n_2 <- length(y_2)

# ---------------------------------------------------------------------
# Build the JOINT (raw) time vector and a SINGLE normalisation
# ---------------------------------------------------------------------

t_raw_joint <- c(t_raw_1, t_raw_2)
t_min       <- min(t_raw_joint)
t_range     <- max(t_raw_joint) - min(t_raw_joint)
t           <- (t_raw_joint - t_min) / t_range
y           <- c(y_1, y_2)
d_idx       <- c(rep(1L, n_1), rep(2L, n_2))
N           <- length(y)
a           <- t_range
```

```{r joint-info, results = "asis"}
cat("- Drop 1: ", n_1, " observations, raw time in [",
    round(min(t_raw_1), 4), ", ", round(max(t_raw_1), 4),
    "] s, initial height = ", round(max(y_1), 4), "\n", sep = "")
cat("- Drop 2: ", n_2, " observations, raw time in [",
    round(min(t_raw_2), 4), ", ", round(max(t_raw_2), 4),
    "] s, initial height = ", round(max(y_2), 4), "\n", sep = "")
cat("- Joint dataset: ", N, " observations, ",
    "t_range = ", round(t_range, 4), " s\n", sep = "")
```

```{r joint-mcmc}
n_iter      <- 10000
burn_in     <- 2000

# init = c(g, h0_1, h0_2, sig2err, alpha, psidelta, k)  (length 7)
init <- c(9.8,
          max(y_1),
          max(y_2),
          0.01, 0.5, 0.5, 0.1)

sigma_props_joint <- c(NA, NA, NA, NA, NA, 0.5, NA)

set.seed(12345)

res_joint <- mcmc_step6_joint(
  y               = y,
  t               = t,
  d_idx           = d_idx,
  n_iter          = n_iter,
  init            = init,
  sigma_proposals = sigma_props_joint,
  g_init          = TRUE,
  h0_init         = FALSE,
  sig2er_init     = FALSE,
  alpha_init      = FALSE,
  psi_init        = FALSE,
  k_init          = FALSE,
  Sigma_theta     = diag(0.5, 3),
  n_burnin        = burn_in,
  seuil           = TRUE,
  s               = 0.3
)

result_joint <- list(
  ball      = "Blue_Basketball",
  sheet_idx = 2,
  n_1       = n_1, n_2 = n_2, N = N,
  t         = t,
  y         = y,
  d_idx     = d_idx,
  t_min     = t_min,
  t_range   = t_range,
  init      = init,
  res       = res_joint
)

save(result_joint,
     file = file.path(out_dir_joint,
                      "joint_seuil_gfix_Blue_Basketball.RData"))
```

## Joint posterior summary

```{r joint-summary, results = "asis"}
theta_mat <- res_joint$theta
summ <- function(x) c(mean = mean(x), median = median(x), sd = sd(x),
                      q025 = unname(quantile(x, 0.025)),
                      q975 = unname(quantile(x, 0.975)))
tab_joint <- rbind(
  g           = summ(theta_mat[, "g"]),
  h0_1        = summ(theta_mat[, "h0_1"]),
  h0_2        = summ(theta_mat[, "h0_2"]),
  sigma_sq    = summ(theta_mat[, "sigma_sq_err"]),
  alpha       = summ(theta_mat[, "alpha"]),
  psi_delta   = summ(theta_mat[, "psi_delta"]),
  k           = summ(theta_mat[, "k"])
)
print(knitr::kable(round(tab_joint, 4),
                   format  = "html",
                   caption = "Joint posterior summary (Blue Basketball)"))
```


```{r joint-figure, fig.width = 14, fig.height = 9}
delta_mat <- res_joint$delta
zeta_mat  <- res_joint$zeta

g_chain     <- theta_mat[, "g"]
h0_1_chain  <- theta_mat[, "h0_1"]
h0_2_chain  <- theta_mat[, "h0_2"]
sigma_chain <- theta_mat[, "sigma_sq_err"]
alpha_chain <- theta_mat[, "alpha"]
psi_chain   <- theta_mat[, "psi_delta"]
k_chain     <- theta_mat[, "k"]

idx1 <- which(d_idx == 1)
idx2 <- which(d_idx == 2)

# alpha density
df_alpha <- data.frame(alpha = alpha_chain)
p_alpha <- ggplot(df_alpha, aes(x = alpha)) +
  geom_density(fill = "#4CCDC9", color = "lightseagreen",
               alpha = 0.6, linewidth = 1) +
  labs(x = expression(alpha), y = "Density",
       title = "Posterior density of alpha (joint, Blue_Basketball)") +
  theme_minimal(base_size = 13) +
  theme(plot.title = element_text(hjust = 0.5))

# delta(x_i), one panel per drop
make_delta_panel <- function(idx, drop_label) {
  d_mat  <- delta_mat[, idx, drop = FALSE]
  d_mean <- colMeans(d_mat)
  d_ci   <- t(apply(d_mat, 2, quantile, probs = c(0.025, 0.975)))
  df <- data.frame(
    i      = seq_along(idx),
    d_mean = d_mean,
    d_low  = d_ci[, 1],
    d_high = d_ci[, 2]
  )
  ggplot(df, aes(x = i, y = d_mean)) +
    geom_ribbon(aes(ymin = d_low, ymax = d_high),
                fill = "lightseagreen", alpha = 0.25) +
    geom_line(color = "lightseagreen", linewidth = 0.9) +
    geom_hline(yintercept = 0, linetype = "dashed") +
    labs(x = "Index i", y = expression(delta(x[i])),
         title = paste0("delta(x_i) — ", drop_label)) +
    theme_minimal(base_size = 12) +
    theme(plot.title = element_text(hjust = 0.5))
}
p_delta_1 <- make_delta_panel(idx1, "drop 1")
p_delta_2 <- make_delta_panel(idx2, "drop 2")

# inclusion probabilities, one panel per drop
make_phat_panel <- function(idx, drop_label) {
  p_hat <- colMeans(zeta_mat[, idx, drop = FALSE] == 2)
  df <- data.frame(i = seq_along(idx), p_hat = p_hat)
  ggplot(df, aes(x = i, y = p_hat)) +
    geom_col(fill = "lightseagreen") +
    ylim(0, 1) +
    labs(x = "Index i",
         y = expression(hat(p)[i] == P(zeta[i] == 1 ~ "|" ~ Y)),
         title = paste0("Inclusion probabilities — ", drop_label)) +
    theme_minimal(base_size = 12) +
    theme(plot.title = element_text(hjust = 0.5))
}
p_zeta_1 <- make_phat_panel(idx1, "drop 1")
p_zeta_2 <- make_phat_panel(idx2, "drop 2")

# boxplot grid
make_box <- function(x, ylab) {
  df <- data.frame(val = x)
  ggplot(df, aes(x = 1, y = val)) +
    geom_boxplot(fill = "lightseagreen", width = 0.25, outlier.size = 1) +
    labs(x = NULL, y = ylab) +
    theme_minimal(base_size = 11) +
    theme(axis.text.x  = element_blank(),
          axis.ticks.x = element_blank(),
          plot.margin  = margin(3, 6, 3, 6))
}

p_g     <- make_box(g_chain,     "g")
p_h0_1  <- make_box(h0_1_chain,  expression(h[0]^{(1)}))
p_h0_2  <- make_box(h0_2_chain,  expression(h[0]^{(2)}))
p_sig   <- make_box(sigma_chain, expression(lambda^2))
p_a_box <- make_box(alpha_chain, expression(alpha))
p_psi   <- make_box(psi_chain,   expression(gamma[delta]))
p_k     <- make_box(k_chain,     "k")

left_grid <- (p_g    | p_h0_1) /
             (p_h0_2 | p_sig)  /
             (p_a_box| p_psi)  /
             (p_k    | plot_spacer())

right_col <- p_alpha /
             (p_delta_1 | p_delta_2) /
             (p_zeta_1  | p_zeta_2) +
             patchwork::plot_layout(heights = c(1, 1.2, 1))

fig_joint <- left_grid | right_col
fig_joint <- fig_joint + patchwork::plot_layout(widths = c(0.9, 1.6))

print(fig_joint)
```


